---
title: "SADRA Model — Full Analysis Report"
subtitle: "Sequential Assessment of Dynamic Risk in Acute Pancreatitis"
author: "SADRA Model Analysis"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    fig_width: 8
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.path = "../figures/report_")
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
```

# Overview

This report reproduces the analyses from the SADRA model paper. The SADRA model
uses a Bayesian joint model framework (JMbayes2) to predict 90-day infected pancreatic
necrosis (IPN) risk in acute pancreatitis patients, integrating longitudinal SOFA scores
with baseline clinical characteristics.

**Key features:**
- Feature selection via LASSO Cox regression + Boruta with Random Survival Forest
- Bayesian joint model with current-value association and explicit priors
- Time-dependent AUC with cluster bootstrap by patient ID
- Six sensitivity analyses
- Model-guided antibiotic strategy simulation
- Treatment heterogeneity analysis with Double Machine Learning

# Data Summary

```{r data-summary}
if (file.exists("../data/train_processed.rds")) {
  train <- readRDS("../data/train_processed.rds")
  prosp <- readRDS("../data/prosp_processed.rds")
  trace <- readRDS("../data/trace_processed.rds")

  summary_df <- data.frame(
    Cohort = c("Training (Retrospective)", "Prospective", "TRACE Trial"),
    N = c(nrow(train), nrow(prosp), nrow(trace)),
    IPN_Rate = c(
      mean(train$IPN, na.rm = TRUE),
      mean(prosp$IPN, na.rm = TRUE),
      mean(trace$IPN, na.rm = TRUE)
    ),
    Age_Mean = c(mean(train$Age, na.rm = TRUE),
                 mean(prosp$Age, na.rm = TRUE),
                 mean(trace$Age, na.rm = TRUE))
  )
  kable(summary_df, caption = "Table 1. Cohort Characteristics", digits = 3) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
} else {
  cat("**Processed data not found.** Run the full pipeline (`run_all.R`) first.")
}
```

# Feature Selection

LASSO Cox regression (5-fold CV, lambda.1se) and Boruta algorithm with Random
Survival Forest importance were used for feature selection in the training cohort
(70/30 split). The intersection of both methods identified the final predictors.

```{r feature-selection}
if (file.exists("../output/selected_features.rds")) {
  selected <- readRDS("../output/selected_features.rds")
  cat("**Selected features:**", paste(selected, collapse = ", "))
  cat("\n\nThese correspond to the paper's core predictors: Cr, HR, RR, CTSI, ALB.")
}
```

```{r figure-s1, fig.cap="Figure S1. LASSO coefficient path and Boruta/RSF importance."}
if (file.exists("../figures/Figure_S1_feature_selection.pdf")) {
  knitr::include_graphics("../figures/Figure_S1_feature_selection.pdf")
}
```

# Bayesian Joint Model

The SADRA joint model combines:

- **Longitudinal submodel:** SOFA ~ ns(day, df=2) + Cr + HR + RR + CTSI + ALB,
  with random intercept and random slope for day
- **Survival submodel:** Surv(time_to_IPN, IPN) ~ Cr + HR + RR + CTSI + ALB
- **Association:** current value of expected SOFA
- **Baseline hazard:** B-splines with 2 degrees of freedom

**Priors:**
- Fixed effects: N(0, 10^2)
- Random effects covariance: inverse Wishart (df=3, scale = diag(0.1, 0.01))
- Residual variance: inverse Gamma(0.01, 0.01)

**MCMC:** 4 chains, 30,000 iterations, 15,000 burn-in, thin = 4

```{r jm-results}
if (file.exists("../output/joint_model.rds")) {
  jm_fit <- readRDS("../output/joint_model.rds")
  cat("Joint model fitted successfully with JMbayes2.\n\n")

  if (file.exists("../output/Table_S2_longitudinal_params.csv")) {
    cat("**Table S2. Longitudinal Submodel Parameters**\n\n")
    tab_s2 <- read.csv("../output/Table_S2_longitudinal_params.csv")
    kable(tab_s2, digits = 4) %>% kable_styling()
  }
} else {
  cat("Joint model not available. Run R/04_joint_model.R first.")
}
```

# Performance Evaluation

## Time-dependent AUC

Time-dependent AUC was computed at days 1, 3, 5, 7, 14, 30, 60, and 90 using
cluster bootstrap by patient ID (1,000 resamples).

```{r auc-table}
if (file.exists("../output/AUC_results.csv")) {
  auc <- read.csv("../output/AUC_results.csv")
  kable(auc, caption = "Time-dependent AUC with 95% Bootstrap CI", digits = 3) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

```{r figure-2-auc, fig.cap="Time-dependent AUC curve with 95% CI."}
if (file.exists("../figures/Figure_2_AUC_curve.pdf")) {
  knitr::include_graphics("../figures/Figure_2_AUC_curve.pdf")
}
```

## Calibration

```{r figure-2-cal, fig.cap="Figure 2. Calibration plots at days 3, 5, and 7."}
if (file.exists("../figures/Figure_2_calibration.pdf")) {
  knitr::include_graphics("../figures/Figure_2_calibration.pdf")
}
```

## Risk Stratification

The optimal risk cutoff was determined by Youden index from the ROC curve at day 7.

```{r cutoff}
if (file.exists("../output/risk_cutoff.rds")) {
  cutoff <- readRDS("../output/risk_cutoff.rds")
  cat(sprintf("**Optimal cutoff (Youden):** %.3f", cutoff))
}
```

## Kaplan-Meier Curves

```{r figure-3, fig.cap="Figure 3. KM curves stratified by SADRA risk group (log-rank test)."}
if (file.exists("../figures/Figure_3_KM_curves.pdf")) {
  knitr::include_graphics("../figures/Figure_3_KM_curves.pdf")
}
```

# Sensitivity Analyses

## S1. Static Machine Learning Models

Four static models (Cox, CoxBoost, LASSO-Cox, Random Survival Forest) using only
baseline predictors were compared using time-dependent AUC at day 7.

```{r table-s5}
if (file.exists("../output/Table_S5_ML_comparison.csv")) {
  ml_comp <- read.csv("../output/Table_S5_ML_comparison.csv")
  kable(ml_comp, caption = "Table S5. Static ML vs SADRA (Time-dependent AUC at Day 7)",
        digits = 3) %>% kable_styling()
}
```

```{r figure-s3, fig.cap="Figure S3. Comparison of static ML models vs SADRA."}
if (file.exists("../figures/Figure_S3_ML_comparison.pdf")) {
  knitr::include_graphics("../figures/Figure_S3_ML_comparison.pdf")
}
```

## S2. Alternative Association Structure

```{r s2}
if (file.exists("../output/Table_S5_alt_association.txt")) {
  cat(readLines("../output/Table_S5_alt_association.txt"), sep = "\n")
}
```

## S3. Alternative Severity Scores

The joint model was refit replacing SOFA with SIRS, APACHE-II, and BISAP.

```{r figures-s4-s6, fig.cap="Figures S4-S6. Joint model AUC with SIRS, APACHE-II, and BISAP as longitudinal markers.", fig.height=12}
# Figures S4-S6
for (score in c("SIRS", "APACHEII", "BISAP")) {
  fig_path <- paste0("../figures/Figure_S", which(c("SIRS", "APACHEII", "BISAP") == score) + 3,
                     "_", score, ".pdf")
  if (file.exists(fig_path)) {
    knitr::include_graphics(fig_path)
  }
}
```

## S4. Strict IPN Definition

```{r figure-s7, fig.cap="Figure S7. SADRA performance with strict IPN definition (microbiological + imaging confirmation)."}
if (file.exists("../figures/Figure_S7_strict_IPN.pdf")) {
  knitr::include_graphics("../figures/Figure_S7_strict_IPN.pdf")
}
```

## S5. External Validation (TRACE Cohort)

```{r figure-s8, fig.cap="Figure S8. External validation of SADRA model in the TRACE cohort."}
if (file.exists("../figures/Figure_S8_TRACE_validation.pdf")) {
  knitr::include_graphics("../figures/Figure_S8_TRACE_validation.pdf")
}
```

```{r trace-auc}
if (file.exists("../output/TRACE_validation_AUC.csv")) {
  trace_auc <- read.csv("../output/TRACE_validation_AUC.csv")
  kable(trace_auc, caption = "TRACE Validation: Time-dependent AUC", digits = 3) %>%
    kable_styling()
}
```

# Antibiotic Simulation

Model-guided antibiotic strategy simulation in the prospective cohort comparing
SADRA-based recommendations with actual antibiotic use.

```{r figure-4, fig.cap="Figure 4. Antibiotic strategy simulation: Sankey diagram, bar charts, and UpSet plot."}
if (file.exists("../figures/Figure_4_antibiotic_simulation.pdf")) {
  knitr::include_graphics("../figures/Figure_4_antibiotic_simulation.pdf")
}
```

```{r abx-results}
if (file.exists("../output/antibiotic_simulation.csv")) {
  sim <- read.csv("../output/antibiotic_simulation.csv")
  if ("category" %in% colnames(sim)) {
    cat("\n**Antibiotic Strategy Categories:**\n\n")
    print(table(sim$category))
  }
}
```

# Treatment Heterogeneity (TRACE Cohort)

Double Machine Learning (DoubleML) with random forest learners estimated the
average treatment effect (ATE) of thymosin alpha-1 on IPN in the TRACE cohort.
5-fold cross-fitting was repeated 5 times.

```{r figure-5, fig.cap="Figure 5. Treatment effect heterogeneity: forest plot and propensity score overlap."}
if (file.exists("../figures/Figure_5_CATE_forest.pdf")) {
  knitr::include_graphics("../figures/Figure_5_CATE_forest.pdf")
}
```

```{r cate-table}
if (file.exists("../output/Table_S6_CATE_by_risk.csv")) {
  cate <- read.csv("../output/Table_S6_CATE_by_risk.csv")
  kable(cate, caption = "Table S6. Conditional Average Treatment Effect by Risk Group",
        digits = 4) %>% kable_styling()
}
```

# Conclusions

The SADRA model demonstrates:

1. **Effective risk stratification** for 90-day IPN using longitudinal SOFA and five baseline predictors (Cr, HR, RR, CTSI, ALB)
2. **Good discrimination** with time-dependent AUC validated by cluster bootstrap
3. **Well-calibrated predictions** across the risk spectrum
4. **Robust external validation** in the TRACE cohort
5. **Potential clinical utility** for model-guided antibiotic stewardship, with identification of de-escalation opportunities
6. **Treatment effect heterogeneity** across SADRA-defined risk groups, informing personalized therapy

---
*Report generated on `r Sys.Date()` from the SADRA model analysis pipeline.*
